---
title: "Wildcat Strikes"
output:
  html_document:
    df_print: paged
---


```{r}
library(dplyr)

```
> Reading the dataset into a dataframe

```{r}
wildcat<-read.csv('/Users/apple/Gdrive/Term1/SA3/Assignment/wildcat.csv',header=TRUE)
```
> Question (a) - Exploration of Dataset:

+ Grievances is a discrete variable (numeric count data). This is highly skewed to the right
+ Rotate, Union are binary variable to describe if Rotation is allowed & if the mine is registered as a Union
+ Workforce is a discrete variable (numeric count data). This variable is highly skewed to the right
+ Log.workforce is a continous and normalized variable
+ Wildcat.Strikes is a discrete variable (numeric count data). This variable is highly skewed to the right


```{r}
summary(wildcat)
```
```{r}
#install.packages('e1071')
library(e1071)
skewness(wildcat$Grievances)
hist(wildcat$Grievances)
hist(wildcat$Workforce)
hist(wildcat$Wildcat.strikes)
```
```{r}
boxplot(wildcat$Log.workforce)
```
> Question (a):

* **There is one mine (Mine # 163), where is there is excessive number of strikes. It is 1.5 times, the # of strikes to the second best mine ** 

```{r}
barplot(wildcat$Wildcat.strikes)
```
> Question (b):

* Total number of wildcat strikes are higher in mines which have unions. To state there are 389 wilcat strikes in Unionized mines, and only 20 strikes in non-Union mines.
* This could also be confirmed that in the non-Unionized mines - the average number of wildcat strikes is 0.345, where as in the Unionized it is 3.70
* While, the total number of strikes in non-Rotatable mines is higher than that of the Rotatable mines, the difference is not higher
```{r}
count(wildcat,wildcat$Union)
aggregate(Wildcat.strikes~Union, data=wildcat, mean)
wildcat %>% group_by(Union) %>% summarise_all(sum) %>% select(Union,Wildcat.strikes)
wildcat %>% group_by(Rotate) %>% summarise_all(sum) %>% select(Rotate,Wildcat.strikes)
aggregate(Wildcat.strikes~Rotate, data=wildcat, mean)

plot(wildcat$Workforce,wildcat$Wildcat.strikes)

```

```{r}
model1<-glm(Wildcat.strikes~ Grievances + Rotate + Union + Workforce,family = poisson(link="log"),data=wildcat)
summary(model1)
```
> Question (c):

* Model has been fit with Poisson distribution
* From the initial analysis, the model seems to be fitting decently. However there is an overdispersion (proved in the subsequent questions)
* Chi-square value and the plots have been drawn to derive the conclusion given below.

```{r}
model2<-glm(Wildcat.strikes~ Grievances + Rotate + Union + Log.workforce,family = poisson(link="log"),data=wildcat)
summary(model2)
```

```{r}
library(MASS)
library(car)
StudentResidualPlot <- function(model, comment){
  stu.resid <- studres(model)  
  hist(stu.resid, freq=FALSE,     
       main=comment) 
  xfit<-seq(min(stu.resid),max(stu.resid),length=80)  
  yfit<-dnorm(xfit)  
  lines(xfit, yfit)
  
}

CustomResidualPlot<- function(model){
  residualPlot(model, id.n=5)
  residualPlots(model,id.n=3)
}

DeletionDiagonsitics<- function(model, cutoff){
  influencePlot(model,id.n=3) # A user friendly representation of the above
  plot(model, which=4, cook.levels=cutoff)
}
```

```{r}
#StudentResidualPlot(model2,'Model2')
#CustomResidualPlot(model2)
#DeletionDiagonsitics(model2,2)
#residualPlot(model2)
#influencePlot(model2,id.n=3)
par(mfrow=c(2,2))
plot(model2)

```

```{r}
pchisq(model2$deviance, df=model2$df.residual, lower.tail=FALSE)


```
```{r}
#install.packages('pscl')
```
```{r}
model3<-glm(Wildcat.strikes~ Grievances + Rotate + Union + Log.workforce,family = quasipoisson(link="log"),data=wildcat)
summary(model3)
```
```{r}
pchisq(model3$deviance, df=model3$df.residual, lower.tail=FALSE)

```
```{r}
#install.packages('AER')
```
> Question (d) : Overdispersion

* There is a very clear over-dispersion from the below statistic, where the variance is approximately 3 times to that of the mean

```{r}
library(AER)
dispersiontest(model2)
```

> Question (e): Other models 

* Trying with other models - Negative Binomial & Zero Inflated.
* negbin and zero infalted provide better fit to the data than poission, as explained below


```{r}
model4<-glm.nb(Wildcat.strikes~ Grievances + Rotate + Union + Log.workforce,data=wildcat)
summary(model4)

```

```{r}
with(wildcat, cbind(res.deviance=model4$deviance, df=model4$df.residual, p=pchisq(model4$deviance, model4$df.residual, lower.tail=FALSE)))

```
```{r}
#ZINB=zeroinfl(Wildcat.strikes~ Grievances + Rotate + Union + Log.workforce, data = na.omit(wildcat), dist = "negbin")
#summary(ZINB)
```

```{r}
par(mfrow=c(2,2))
plot(model4)
```

